Hierarchical Time Series Forecasting in Emergency Medical Services

### 2023-12-16 manuscript/main.tex
second revision finalised

\begin{keyword}
    healthcare \sep emergency services \sep forecast
reconciliation \sep ambulance demand \sep 
    regression
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, enhanced, frame hidden]}{\end{tcolorbox}}\fi

@@

Demand forecasts at EMS are typically required at multiple levels of an
organization to inform various planning and decision-making processes
\citep{hulshof2012taxonomic}. There are some planning processes at the
national level (strategic and long-term) such as workforce resource
planning and budgeting; sub-national, regional, or healthcare level
(tactical and medium-term) such as temporary capacity expansions,
resource sharing; and hospital or station level (operational and
short-term) such as planning rosters for staff and ambulance deployment.
Demand forecasts might also be required at different levels for a
specific area of interest such as the nature of demand or the priority
level. Moreover, the time series data in EMS has an inherent
hierarchical and grouped structure to support such forecasting
requirements. Demand for emergency medical services at the national
level can be disaggregated in a geographical hierarchy into
sub-national, regions, health boards, and stations/hospitals, or divided
into groups such as the nature of incidents or demand priority.
\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchies are necessary for effective decision-making in EMS. For example, control area EMS forecasts can inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals. At the lower levels, hospitals or ambulance stations could use such forecasts to plan for staffing and resource allocation, ambulance dispatching, staff-to-shift assignment, staff rescheduling based on the anticipated volume and priority and nature of incidents. Additionally, generating forecasts at lower levels could potentially improve the accuracy of the high-level forecasts, by providing more detailed information on the nature and priority of incidents. This could help to identify patterns in demand that may not be apparent at the higher level. Therefore, employing forecasting techniques that consider the hierarchical and/or grouped patterns of time series in EMS aligns naturally, offering the possibility to enhance forecast accuracy and facilitate coordination.}

@@

Despite a large number of studies dedicated to forecasting for EMS
\citep{mingliterature2022, gul2020exhaustive, ibrahim2016modeling, wargon2009systematic},
the hierarchical data structure has been largely ignored, and the main
focus has been on producing independent (base) forecasts at a single
level. Generating independent forecasts can result in a lack of
consistency and coordination, and therefore leads to less effective
planning and decision making.
\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area. If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area. Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others. Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly. Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}

@@

In this paper, we address this gap by investigating the application of
hierarchical forecasting approaches in the EMS using daily time series
of attended incidents from 2015 to 2019 in a major ambulance service in
Great Britain. The data has hierarchical and grouped structures, with
hierarchies at the national, control (i.e.~sub-national), and health
board (i.e.~regional) levels, as well as groups by priority and nature
of incidents. We produce consistent point forecasts and forecast
distributions for all levels, which is critical for an effective
planning and associated risk management. We compare the point and
probabilistic forecast accuracy of the independent forecasts, bottom-up
and optimal reconciliation approaches. We first generate
independent/base forecasts using Exponential Smoothing State Space
(ETS), Poisson regression using Generalized Linear Model (GLM) and
tscount (TSGLM), a simple empirical distribution and an ensemble method,
followed by applying bottom-up and optimal reconciliation approaches.
Forecast performance is assessed by the Mean Absolute Scaled Error
(MASE) and Mean Squared Scaled Error (MSSE) for point forecasts and
Continuous Ranked Probability Scores (CRPS) for the probabilistic
forecasts. This paper complies with reproducibility principles
\citep{stodden2013best, boylan2015reproducibility}. We provide the R
codes for the proposed models and benchmarks. Therefore, they can be
applied to any healthcare service (e.g., emergency department, primary
or social care) subject to the time series having a hierarchical and/or
grouped structure.

@@

\citet{sasaki2010using} develop a multivariable regression model to
estimate future EMS demands. In addition to the historical demand, the
population census for different age groups and counts of the number of
companies employing more than five people are included in the
regression. The census variables describe groups who are more likely to
need an ambulance. A stepwise ordinary least squares regression analysis
is used for estimating the parameter and generating forecasts. The only
performance measure reported in this study is \(R^2\), which is not an
effective measure of forecast accuracy \citep[p457]{armstrong01}. The
research design of this study is not rigorous and the study is not
reproducible. \citet{vile2012predicting} explore using a Singular
Spectrum Analysis (SSA) method to generate forecasts of the EMS demand
at the national level for 7-day, 14-day, 21-day, and 28-day forecast
horizons using data provided by an ambulance service in Great Britain.
The performance of this approach is compared to Auto-Regressive
Integrated Moving Average (ARIMA) and Holt-Winters time series methods
using Root Mean Squared Error (RMSE). They concluded that point
forecasts generated by SSA are more accurate for longer-term, but that
ARIMA and Holt-Winters performance is superior for shorter-term
horizons. \citet{vile2016time} further develop a decision support system
to integrate forecasts generated by SSA. However, the study does not
compare and contrast the performance of forecasting methods based on
utility measures such as cost, resource utilization and response time.
The tool contains options that allow generating forecasts at various
levels of granularity; however, it ignores the structure of the
hierarchical and grouped relationships, preventing aligned decision
making and coordination. \citet{al2021empirical}
\textcolor{teal}{utilises data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand. The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand. Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE). The findings indicate that ARIMA performs the best in predicting weekly and monthly demand. However, when it comes to long-term demand, the SSA method proves to be the most effective.}
\citet{ibrahim2016modeling}
\textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2). Call centre data were utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days. The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval. The findings indicate that the ME consistently produces the most accurate point forecasts. On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead. For a two-week leading period, MU shows better coverage probability.}

@@

\resizebox{\linewidth}{!}{
\begin{tabular}{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}lll}
\toprule
Reference & Year & Variable & Horizon & Method & Metric & Probabilistic & Reconciliation & Reproducibility\\
\midrule
Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES & YES\\
Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO & No\\
Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO & No\\
Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO & No\\
Ibrahim et al. & 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO & No\\
Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO & No\\
Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO & No\\
\bottomrule
\end{tabular}}
\endgroup{}
\end{table}
\end{landscape}

@@

\subcaption{\label{fig-hierarchy-1}Hierarchical structure: Attended
incidents in the whole country are disaggregated into 3 control areas
and then into 7 different healthboards.}
\end{minipage}%
%
\begin{minipage}[t]{0.02\linewidth}

@@

\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by}
\citet{mstl}.
\textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components. The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component. The strength of trend is defined as:}
\[F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)\]
\textcolor{teal}{For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component. Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small. But for data with little or no trend, the two variances should be approximately the same.}

\textcolor{teal}{The strength of seasonality is defined similarly:}
\[F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).\]
\textcolor{teal}{series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).}

@@

In addition to displaying the trend and seasonality features, we also
visualize a few time series at various levels of aggregation.
Figure~\ref{fig-dataviz2} reveals different information such as trend,
seasonality, and noise. For example, some series depict seasonality and
trend, whereas some other series report low volume of attended incidents
and entropy, making them more volatile and difficult to forecast. At the
level on nature of incidents combined with categories of other levels,
there are many series that contain zeros with low counts. As such, the
data set represents a diverse set of daily time series patterns.

@@

\textbf{Stationary:} We start with a simple forecasting approach,
assuming that the future days will be similar to past days. We use the
empirical distribution of the past daily attended incidents to create
the forecast distribution of future attended incidents.
\textcolor{teal}{We have chosen this "stationary" method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users. Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast. Additionally, the stationary method has shown surprisingly high accuracy. Hence, any forecasting approach that can offer superior results compared to the stationary method would validate its practical use, otherwise there is no necessity for employing more complex methods.}

@@

\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series. Specifically, we utilize the `ETS()` function in the fable package of R, which employs Akaike's Information Criterion (AIC) to identify the optimal model for each time series. Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series. Consequently, the automated algorithm selects the best model based on the unique characteristics of each individual time series. As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}

@@

\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
\begin{align*}
  y_t &\sim \text{Poisson}(\lambda_t) \\
  \text{where}\qquad
  \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
\end{align*}
\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the mean of the Poisson distribution. In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise). The Fourier terms are as defined in}
\citet[Section 7.4]{hyndman2021forecasting}
\textcolor{teal}{This model takes account of weekly seasonality and annual seasonality. Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand. There is no reason, for example, for incidents to occur more at some times of the month than others.}

@@

\textbf{Poisson Regression using tscount (TSGLM):} We also consider
another Poisson regression model that takes into account serial
dependence. This model captures the short range serial dependence by
including autoregressive terms, in addition to the same covariates that
were used in the GLM model. To distinguish this from the previous GLM
model, we will refer to this model as \texttt{TSGLM}.

\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional autoregressive component accounting for serial correlation. The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points, after accounting for the exogenous covariates. Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence. TSGLM can be written as}
\begin{align*}
  y_t &\sim \text{Poisson}(\lambda_t) \\
  \text{where}\qquad
  \log(\lambda_t) &= (\bm{y}_{t-k}' , \bm{x}_t')\bm{\beta},
\end{align*}
\textcolor{teal}{and ${y}_{t-k}$ is a vector of $k$ lagged values. The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e. past values) of the ambulance demand in the model. This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}

\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R}
\citep{JSSv082i05}
\textcolor{teal}{to model the attended incidents. Again, the logarithmic link function is used to ensure that the mean of the Poisson distribution is always positive.}

\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution}
\citep[p156--158]{feller1991introduction}.
\textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context. To account for changes over time, we incorporate trend and seasonality covariates, as well as public holiday effects, allowing the mean of the Poisson distribution to vary. However, it is important to note that if there are additional factors influencing the mean of the Poisson distribution, which are not accounted for in our model, we might observe over-dispersion or under-dispersion in the data.}

\textbf{Ensemble method:}
\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by combining the forecasts, rather than relying on separate forecasts generated by each individual method}
\citep{clemen1989combining}.
\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Stationary, ETS, GLM and TSGLM models using a simple average to form a mixture distribution}
\citep{combinations}.

\textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in}
\citet{panagiotelis2023probabilistic}.
\textcolor{teal}{This involves simulating 1000 future sample paths from each of the models, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series. In this way, we can generate an empirical distribution of forecasts for each model. The ensemble forecast distribution is a simple mixture of these empirical distributions.}

\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models, or to promote a particular model class. Instead, we have developed a flexible framework that can accommodate any forecasting models. Our primary objective is to demonstrate its practicality and effectiveness in integrating base forecasts from any model and generating coherent forecasts within a hierarchical structure.}

@@

\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by $j$, representing the number of time periods ahead we are predicting. In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}

@@

\textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations. It measures how well the predicted probabilities match the observations. On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast. A model is well-calibrated if the predicted probabilities match the distribution of the observations, and it is sharp if it is confident in its predictions. The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated. Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.}

\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores}
\citep[Section 5.9]{winkler1972decision, hyndman2021forecasting}
\textcolor{teal}{or percentile scores} \citep[Section
5.9]{hyndman2021forecasting},
\textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.}
\textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score. Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g. 95\%) quantiles. In such cases, a percentile score becomes useful in meeting this specific requirement.}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}

>>>>

\begin{keyword}
    healthcare \sep emergency services \sep forecast
reconciliation \sep ambulance demand \sep 
    regression
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

Demand forecasts at EMS are typically required at multiple levels of an
organization to inform various planning and decision-making processes
\citep{hulshof2012taxonomic}. There are some planning processes at the
national level (strategic and long-term) such as workforce resource
planning and budgeting; sub-national, regional, or healthcare level
(tactical and medium-term) such as temporary capacity expansions,
resource sharing; and hospital or station level (operational and
short-term) such as planning rosters for staff and ambulance deployment.
Demand forecasts might also be required at different levels for a
specific area of interest such as the nature of demand or the priority
level. Moreover, the time series data in EMS has an inherent
hierarchical and grouped structure to support such forecasting
requirements. Demand for emergency medical services at the national
level can be disaggregated in a geographical hierarchy into
sub-national, regions, health boards, and stations/hospitals, or divided
into groups such as the nature of incidents or demand priority.
Forecasts produced at both higher and lower levels of hierarchies are
necessary for effective decision-making in EMS. For example, control
area EMS forecasts can inform strategic decisions about how to allocate
limited resources to lower levels, such as health boards and
stations/hospitals. At the lower levels, hospitals or ambulance stations
could use such forecasts to plan for staffing and resource allocation,
ambulance dispatching, staff-to-shift assignment, staff rescheduling
based on the anticipated volume and priority and nature of incidents.
Additionally, generating forecasts at lower levels could potentially
improve the accuracy of the high-level forecasts, by providing more
detailed information on the nature and priority of incidents. This could
help to identify patterns in demand that may not be apparent at the
higher level. Therefore, employing forecasting techniques that consider
the hierarchical and/or grouped patterns of time series in EMS aligns
naturally, offering the possibility to enhance forecast accuracy and
facilitate coordination.

@@

Despite a large number of studies dedicated to forecasting for EMS
\citep{mingliterature2022, gul2020exhaustive, ibrahim2016modeling, wargon2009systematic},
the hierarchical data structure has been largely ignored, and the main
focus has been on producing independent (base) forecasts at a single
level. Generating independent forecasts can result in a lack of
consistency and coordination, and therefore leads to less effective
planning and decision making.

@@

In this paper, we address this gap by investigating the application of
hierarchical forecasting approaches in the EMS using daily time series
of attended incidents from 2015 to 2019 in a major ambulance service in
Great Britain. The data has hierarchical and grouped structures, with
hierarchies at the national, control (i.e.~sub-national), and health
board (i.e.~regional) levels, as well as groups by priority and nature
of incidents. We produce consistent point forecasts and forecast
distributions for all levels over 84 days horizon, which is critical for
an effective planning and associated risk management. We compare the
point and probabilistic forecast accuracy of the independent forecasts,
bottom-up and optimal reconciliation approaches. We first generate
independent/base forecasts using Exponential Smoothing State Space
(ETS), Poisson regression using Generalized Linear Model (GLM) and
tscount (TSGLM), a simple empirical distribution and an ensemble method,
followed by applying bottom-up and optimal reconciliation approaches.
Forecast performance is assessed by the Mean Absolute Scaled Error
(MASE) and Mean Squared Scaled Error (MSSE) for point forecasts and
Continuous Ranked Probability Scores (CRPS) for the probabilistic
forecasts. This paper complies with reproducibility principles
\citep{stodden2013best, boylan2015reproducibility}. We provide the R
codes for the proposed models and benchmarks. Therefore, they can be
applied to any healthcare service (e.g., emergency department, primary
or social care) subject to the time series having a hierarchical and/or
grouped structure. While our research focuses on emergency medical
services, it is important to emphasize the suggested framework's
adaptability, which expands its relevance to a variety of service
sectors such as supply chains, tourism, finance, and call centers. Our
approach can be generalized in cases with hierarchically structured
and/or grouped time series data, which is common in many service
sectors.

@@

\citet{sasaki2010using} develop a multivariable regression model to
estimate future EMS demands. In addition to the historical demand, the
population census for different age groups and counts of the number of
companies employing more than five people are included in the
regression. The census variables describe groups who are more likely to
need an ambulance. A stepwise ordinary least squares regression analysis
is used for estimating the parameter and generating forecasts. The only
performance measure reported in this study is \(R^2\), which is not an
effective measure of forecast accuracy \citep[p457]{armstrong01}. The
research design of this study is not rigorous and the study is not
reproducible. \citet{vile2012predicting} explore using a Singular
Spectrum Analysis (SSA) method to generate forecasts of the EMS demand
at the national level for 7-day, 14-day, 21-day, and 28-day forecast
horizons using data provided by an ambulance service in Great Britain.
The performance of this approach is compared to Auto-Regressive
Integrated Moving Average (ARIMA) and Holt-Winters time series methods
using Root Mean Squared Error (RMSE). They concluded that point
forecasts generated by SSA are more accurate for longer-term, but that
ARIMA and Holt-Winters performance is superior for shorter-term
horizons. \citet{vile2016time} further develop a decision support system
to integrate forecasts generated by SSA. However, the study does not
compare and contrast the performance of forecasting methods based on
utility measures such as cost, resource utilization and response time.
The tool contains options that allow generating forecasts at various
levels of granularity; however, it ignores the structure of the
hierarchical and grouped relationships, preventing aligned decision
making and coordination. \citet{al2021empirical} utilises data from the
Welsh Ambulance Service to explore the forecast accuracy of four
forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and
Singular Spectrum Analysis (SSA) in predicting call volume demand. The
aim is to compare these approaches with the current method across
various planning horizons (7 days, 30 days, and 90 days) for both total
call volume and category-specific demand. Forecast accuracy performance
is evaluated using root mean square error (RMSE) and mean absolute
percentage error (MAPE). The findings indicate that ARIMA performs the
best in predicting weekly and monthly demand. However, when it comes to
long-term demand, the SSA method proves to be the most effective.
\citet{ibrahim2016modeling} conducted a case study to assess the
effectiveness of multiple forecasting methods: the multiplicative
univariate forecasting model (MU), univariate mixed-effects model (ME),
and two variations of bivariate mixed-effects models (BME). Call centre
data were utilized to forecast for periods of 1, 7, and 14 days ahead,
using only a limited dataset of 42 days. The performance of these
forecasting methods was evaluated using two metrics: RMSE for point
forecasts and coverage probability for the 95\% prediction interval. The
findings indicate that the ME consistently produces the most accurate
point forecasts. On the other hand, BME models demonstrate superior
coverage probabilities when forecasting for one day or one week ahead.
For a two-week leading period, MU shows better coverage probability.

@@

\resizebox{\linewidth}{!}{
\begin{tabular}{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}lll}
\toprule
Reference & Year & Variable & Horizon & Method & Metric & Probabilistic & Reconciliation & Reproducibility\\
\midrule
Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES & YES\\
Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO & NO\\
Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO & NO\\
Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO & NO\\
Ibrahim et al. & 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO & NO\\
Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO & NO\\
Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO & NO\\
\bottomrule
\end{tabular}}
\endgroup{}
\end{table}
\end{landscape}

@@

\subcaption{\label{fig-hierarchy-1}Hierarchical structure: Attended
incidents in the whole country are disaggregated into 3 control areas
and then into 7 different healthboards, anonymized using two letters
(e.g.~AB)}
\end{minipage}%
%
\begin{minipage}[t]{0.02\linewidth}

@@

In this paper, the strength of trend and seasonality were calculated
using the ``STL'' (Seasonal and Trend decomposition using Loess)
decomposition method, as described by \citet{mstl}. STL is a widely used
and flexible method for decomposing time series data into trend,
seasonal, and remainder components. The decomposition of a time series
\(y_t\) is written as \(y_t = T_t + S_{t} + R_t\), where \(T_t\) is the
smoothed trend component, \(S_t\) is the seasonal component and \(R_t\)
is a remainder component. The strength of trend is defined as:
\[F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)\]
For strongly trended data, the seasonally adjusted data should have much
more variation than the remainder component. Therefore
Var(\(R_t\))/Var(\(T_t+R_t\)) should be relatively small. But for data
with little or no trend, the two variances should be approximately the
same.

The strength of seasonality is defined similarly:
\[F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).\]
series with seasonal strength \(F_S\), close to 0 exhibits almost no
seasonality, while a series with strong seasonality will have \(F_S\)
close to 1 because Var(\(R_t\)) will be much smaller than
Var(\(S_t+R_t\)).

@@

In addition to displaying the trend and seasonality strength
\citep{hyndman2021forecasting}, we also visualize a few time series at
various levels of aggregation. Figure~\ref{fig-dataviz2} reveals
different information such as trend, seasonality, and noise. For
example, some series depict seasonality and trend, whereas some other
series report low volume of attended incidents and entropy, making them
more volatile and difficult to forecast. At the level on nature of
incidents combined with categories of other levels, there are many
series that contain zeros with low counts. As such, the data set
represents a diverse set of daily time series patterns.

@@

\textbf{Stationary:} We start with a simple forecasting approach,
assuming that the future days will be similar to past days. We use the
empirical distribution of the past daily attended incidents to create
the forecast distribution of future attended incidents. We have chosen
this ``stationary'' method as a benchmark due to its widespread usage
and simplicity, making it easily understandable for users. Forecasts
serve as inputs for various decision-making systems that frequently
employ simulations, wherein it is common to utilize the empirical
distribution of demand as a forecast. Additionally, the stationary
method has shown surprisingly high accuracy. Hence, any forecasting
approach that can offer superior results compared to the stationary
method would validate its practical use, otherwise there is no necessity
for employing more complex methods.

@@

In our study, we use an automated algorithm to determine the suitable
configuration for the trend, seasonality, and error terms in each time
series. Specifically, we utilize the \texttt{ets()} function in the
forecast package of R, which employs Akaike's Information Criterion
(AIC) to identify the optimal model for each time series. Given the
large number of time series we work with (1530), it is impractical to
manually select the appropriate form for each component in every time
series. Consequently, the automated algorithm selects the best model
based on the unique characteristics of each individual time series. As a
result, a combination of additive or multiplicative forms for the
components are employed, depending on the specific attributes of each
time series.

@@

Suppose the time series is denoted by \(y_1,\dots,y_T\), then the
Poisson GLM can be written as \begin{align*}
  y_t &\sim \text{Poisson}(\lambda_t) \\
  \text{where}\qquad
  \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
\end{align*} and \(\bm{x}_t\) is a vector of covariates, \(\beta\) is a
vector of coefficients, and \(\lambda_t\) is the mean of the Poisson
distribution. In our model, these include cubic splines for the time
trend, day-of-week dummy variables (from Monday to Sunday), Fourier
terms to capture the yearly seasonality, dummy variables indicating
public holidays (1 when is a public holiday, 0 otherwise), school
holidays (1 when is a school holiday, 0 otherwise), and Christmas Day (1
when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New
Year's Day, 0 otherwise). The Fourier terms are as defined in
\citet[Section 7.4]{hyndman2021forecasting}. This model takes account of
weekly seasonality and annual seasonality. Monthly seasonality in time
series data is extremely rare, and it does not exist in the ambulance
demand used in this study. There is no reason for occurrences to occur
more frequently at certain times of the month than others.

@@

\textbf{Poisson Regression using tscount (TSGLM):} We also consider
another Poisson regression model that takes into account serial
dependence. This model captures the short-range serial dependence by
including autoregressive terms in addition to the same covariates that
were used in the GLM model. To distinguish this from the previous GLM
model, we will refer to this model as \texttt{TSGLM}.

The Poisson TSGLM is similar to the GLM, with an additional
autoregressive component accounting for serial dependence. The term
serial dependence refers to instances in which the number of incidents
on a current day correlates with the number of incidents on previous
days. \begin{align*}
  y_t &\sim \text{Poisson}(\lambda_t) \\
  \text{where}\qquad
  \log(\lambda_t) &= (\bm{y}_{t-k}' , \bm{x}_t')\bm{\beta},
\end{align*} and \({y}_{t-k}\) is a vector of \(k\) lagged values. The
TSGLM model explicitly accounts for serial dependence by including
lagged values (i.e., past values) of the ambulance demand in the model.
This is important in EMS forecasting because it allows the model to
capture patterns in the data that are dependent on the past values of
the time series, which might not be captured via the predictor
variables.

We use the \texttt{tsglm()} function in the \texttt{tscount} package in
R \citep{JSSv082i05} to model the attended incidents. Again, the
logarithmic link function is used to ensure that the mean of the Poisson
distribution is always positive.

Provided accidents occur independently, they will inherently follow a
Poisson distribution \citep[p156--158]{feller1991introduction}. Hence,
it is reasonable to assume a Poisson distribution in this context. To
account for changes over time, we incorporate trend and seasonality
covariates, as well as public holiday effects, allowing the mean of the
Poisson distribution to vary. However, it is important to note that if
there are additional factors influencing the mean of the Poisson
distribution that are not accounted for in our model, we might observe
over- or under-dispersion in the data.

\textbf{Ensemble method:} Finally, one effective strategy for improving
forecast accuracy includes the simultaneous application of multiple
forecasting methods on a given time series, followed by combining the
forecasts rather than relying on separate forecasts generated by each
individual method \citep{clemen1989combining}. In this paper, we use an
ensemble method that combines the forecasts generated from the
Stationary, ETS, GLM, and TSGLM models using a simple average to form a
mixture distribution \citep{combinations}.

To generate forecast probability distributions using the above methods,
we use a form of bootstrapping, described in
\citet{panagiotelis2023probabilistic}. This involves simulating 1000
future sample paths from each of the models by bootstrapping the model
residuals, taking into account the cross-sectional correlations between
the different aggregated and disaggregated series. In this way, we can
generate an empirical distribution of forecasts for each model. The
ensemble forecast distribution is a simple mixture of these empirical
distributions.

It is important to emphasize that the aim of this study is not to
provide an exhaustive compilation of forecasting models or to promote a
particular model class. Instead, we have developed a flexible framework
that can accommodate any forecasting model. Our primary objective is to
demonstrate its practicality and effectiveness in integrating base
forecasts from any model and generating coherent forecasts within a
hierarchical structure.

@@

The error metrics provided below consider a forecasting horizon denoted
by \(j\), representing the number of time periods ahead we are
predicting. In our study, this forecasting horizon ranges from 1 to 84
days, \(j= 1,2,\dots, 84\).

@@

Calibration refers to the statistical consistency between the
distributional forecasts and the observations. It measures how well the
predicted probabilities match the observations. On the other hand,
sharpness refers to the concentration of the forecast distributions ---
a sharp forecast distribution results in narrow prediction intervals,
indicating high confidence in the forecast. A model is well-calibrated
if the predicted probabilities match the distribution of the
observations, and it is sharp if it is confident in its predictions. The
CRPS rewards sharpness and calibration by assigning lower scores to
forecasts with sharper distributions, and to forecasts that are
well-calibrated. Thus, it is a metric that combines both sharpness and
miscalibration into a single score, making it a useful tool for
evaluating the performance of probabilistic forecasts.

CRPS can be considered an average of all possible Winkler scores
\citep[Section 5.9]{winkler1972decision, hyndman2021forecasting} or
percentile scores \citep[Section 5.9]{hyndman2021forecasting}, and thus
provides an evaluation of all possible prediction intervals or
quantiles. A specific prediction interval could be evaluated using a
Winkler score. Certain situations may also require assessing accuracy
for a particular quantile, such as lower (e.g 5\%) or higher (e.g.~95\%)
quantiles. In such cases, a percentile score becomes useful in meeting
this specific requirement.

@@

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-hstep-1.pdf}
